\vspace{1.0cm} \LARGE \bf MERS-CoV spillover at the camel-human interface

### 2017-11-20 mers-structure.tex
State findings differently. Fixes #30.
Address #29, #19 and reformat for eLife

\documentclass[11pt,oneside,letterpaper]{article}

@@

\usepackage[hidelinks]{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black}

@@

\author[1]{Gytis Dudas}
\author[2]{Luiz Max Carvalho}
\author[2,3]{Andrew Rambaut}
\author[1]{Trevor Bedford}

@@

% \date{\today}

@@

Here, we use an explicit model of metapopulation structure and migration between discrete subpopulations, referred to here as demes \citep{vaughan_efficient_2014}, derived from the structured coalescent \citep{notohara_structured_coalescent_1990}.
Unlike approaches that model host species as a discrete phylogenetic trait of the virus using continuous-time Markov processes (or simpler, parsimony based, approaches) \citep{faria_simultaneously_2013,lycett_h5n8_2016}, population structure models explicitly incorporate contrasts in deme effective population sizes and migration between demes.
By estimating independent coalescence rates for MERS-CoV in humans and camels, as well as migration patterns between the two demes, we show that long-term viral evolution of MERS-CoV occurs exclusively in camels.
Our results suggest that spillover events into humans are seasonal and might be associated with the calving season in camels.
Once human MERS-CoV infections are established, however, we find that MERS-CoV is poor at transmitting between humans.
Using Monte Carlo simulations we show that $R_{0}$ for MERS-CoV circulating in humans is much lower than the epidemic threshold of 1.0 and that correspondingly the virus has been introduced into humans hundreds of times.

@@

\bibliographystyle{mbe}
\bibliography{mers-structure}

>>>>

% \documentclass[11pt,oneside,letterpaper]{article}
\documentclass[9pt,lineno]{elife}

@@

% \usepackage[hidelinks]{hyperref}
% \hypersetup{colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black}

@@

\author[1*]{Gytis Dudas}
\author[2]{Luiz Max Carvalho}
\author[2,3]{Andrew Rambaut}
\author[1]{Trevor Bedford}

@@

\corr{gdudas@fredhutch.org}{GD}

@@

Here, we use an explicit model of metapopulation structure and migration between discrete subpopulations, referred to here as demes \citep{vaughan_efficient_2014}, derived from the structured coalescent \citep{notohara_structured_coalescent_1990}.
Unlike approaches that model host species as a discrete phylogenetic trait of the virus using continuous-time Markov processes (or simpler, parsimony based, approaches) \citep{faria_simultaneously_2013,lycett_h5n8_2016}, population structure models explicitly incorporate contrasts in deme effective population sizes and migration between demes.
By estimating independent coalescence rates for MERS-CoV in humans and camels, as well as migration patterns between the two demes, we show that long-term viral evolution of MERS-CoV occurs exclusively in camels.
Our results suggest that spillover events into humans are seasonal and might be associated with the calving season in camels.
However, we find that MERS-CoV, once introduced into humans, follows transient transmission chains that soon abate.
Using Monte Carlo simulations we show that $R_{0}$ for MERS-CoV circulating in humans is much lower than the epidemic threshold of 1.0 and that correspondingly the virus has been introduced into humans hundreds of times.

@@

% \bibliographystyle{mbe}
\bibliography{mers-structure}
